Multiplication, division and factorization\\of p-adic polynomials

### 2015-12-23 polynomials/muldivfact.tex
Definition of lattice precision

\noindent
{\bf Newton precision.} \todo{Write a one-line introduction.}

\begin{deftn}
A \emph{Newton function of degree $n$} is a convex function 
$f : [0,n] \to \R$ which is piecewise affine and whose epigraph
$\Epi(f)$ have integral extremal points.
\end{deftn}

\begin{rem}
The datum of $f$ is equivalent to that of $\Epi(f)$ and they can
easily represented and manipulated on a computer.
\end{rem}

@@

In a similar fashion, Newton functions can be used to model precision: 
given a Newton function $f$ of degree $n$, we agree that a polynomial 
of degree at most $n$ is given at precision $O(f)$ when, for all $i$,
its $i$-th coefficient is given at precision $O\big(\pi^{\lceil f(i)
\rceil}\big)$ (where $\lceil \cdot \rceil$ is the ceiling function).
In the sequel, we shall write
$O(f) = \sum_{i=0}^n O\big(\pi^{\lceil f(i) \rceil}\big) \cdot X^i$
and use the notation $\sum_{i=0}^n a_i X^i + O(f)$
to refer to a polynomial given at precision $O(f)$.

We observe in particular that Newton precision data are particular
instances of jagged precision. The main interest of using them is the
efficiency.

@@

\noindent
{\bf Lattice precision.}

>>>>

\noindent
{\bf Newton precision.} 
We now move to \emph{Newton precision} data. They can be actually seen 
as particular instances of jagged precision but there exist for them 
better representations and better algorithms.

\begin{deftn}
A \emph{Newton function of degree $n$} is a convex function 
$\varphi : [0,n] \to \R$ which is piecewise affine and whose epigraph
$\Epi(\varphi)$ have integral extremal points.
\end{deftn}

\begin{rem}
The datum of $\varphi$ is equivalent to that of $\Epi(\varphi)$ and they 
can easily represented and manipulated on a computer.
\end{rem}

@@

In a similar fashion, Newton functions can be used to model precision: 
given a Newton function $\varphi$ of degree $n$, we agree that a polynomial 
of degree at most $n$ is given at precision $O(\varphi)$ when, for all $i$,
its $i$-th coefficient is given at precision $O\big(\pi^{\lceil \varphi(i)
\rceil}\big)$ (where $\lceil \cdot \rceil$ is the ceiling function).
In the sequel, we shall write
$O(\varphi) = \sum_{i=0}^n O\big(\pi^{\lceil \varphi(i) \rceil}\big) \cdot X^i$
and use the notation $\sum_{i=0}^n a_i X^i + O(\varphi)$ (where the
coefficients $a_i$ are given by truncated series) to refer to a 
polynomial given at precision $O(\varphi)$.

It is easily checked that if $P$ and $Q$ are two polynomials known at 
precision $O(\varphi_P)$ and $O(\varphi_Q)$ respectively, then $P+Q$ is 
known at precision $O(\varphi_P \oplus \varphi_Q)$ and $PQ$ is known at 
precision $O\big(\varphi_P \odot \NF(Q) \oplus \NF(P) \odot 
\varphi_Q\big)$.

@@

\noindent
{\bf Lattice precision.}
The notion of \emph{lattice precision} was developed in 
\cite{padicprec}. It encompasses the two previous models and has
the decisive advantage of optimality. As a counterpart, it might be
very space-consuming and time-consuming for polynomials of large
degree.

\begin{deftn}
Let $V$ be a finite dimensional vector space over $K$. A lattice
in $V$ is a sub-$W$-module of $V$ generated by a $K$-basis of
$V$.
\end{deftn}

\noindent
We fix an integer $n$. A lattice precision datum for a polynomial of 
degree $n$ is a lattice $H$ lying in the vector space $K_{\leq n}[X]$. 
We shall sometimes denote it $O(H)$ in order to emphasize that it should 
be considered as a precision datum. The notation $P(X) + O(H)$ then 
refers to any polynomial in the $W$-affine space $P(X) + H$. Tracking
lattice precision can be done using differentials as shown in
\cite[Lemma~3.4 and Proposition~3.12]{padicprec}: if $f : K_{\leq n}[X] 
\to K_{\leq m}[X]$ denotes any strictly differentiable function with
surjective differential, under mild assumption on $H$, we have:
$$f(P(X)+H) = f(P(X)) + f'(P(X))(H)$$
where $f'(P(X))$ denotes the differential of $f$ at $P(X)$. The
equality sign reflets the optimality of the method.

As already mentioned, the jagged precision model is a particular case
of the lattice precision. Indeed, a precision of the shape
$\sum_{i=0}^n O(\pi^{N_i}) X^i$ corresponds to the lattice
generated by the elements $\pi^{N_i} X^i$ ($0 \leq i \leq n$).
